---
title: "Merging and Cleaning 2001-11 and 11-17 data"
output: html_notebook
---

Load libraries

```{r message=FALSE}
library(rgdal)
library(spdep)
library(dplyr)
library(lubridate)
library(maptools)
library(rgeos)
```

Read in Spatial Point Files. 2001-10 are in one file and 11-17 are in another.

```{r}
homs_01_10 <- readOGR("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/2001_10/", "hom2001a2010_final")

homs_11_17 <- readOGR("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/2011_17/", "hom2011a2017_06042018" )

```

Rename columns to enable merge

YOD = Year of Death
DOD = Date of Death
HOD = Hour of Death
DOB = Date of Birth

```{r}
#Select specific columns in each
homs_01_10@data <- select(homs_01_10@data, NAME, ANO, DTOBITO, HORAOBITO, DTNASC, IDADE, SEXO, RACACOR, CAUSABAS, LINHAA, LINHAB, LINHAC, LINHAD, LINHAII, nome_novo)

homs_11_17@data <- select(homs_11_17@data, NUMERODO, ANO_OBITO, DTOBITO, HORAOBITO, DTNASC, IDADE, SEXO, RACACOR, CAUSABAS, LINHAA, LINHAB, LINHAC, LINHAD, LINHAII, BAIRRONOVO)

colnames(homs_01_10@data) <- c("possible_ID", "YOD", "DOD", "HOD", "DOB", "Age", "Sex", "Race", "Base_cause", "LineA", "LineB", "LineC", "LineD", "LineII", "Bairro")

colnames(homs_11_17@data) <- c("possible_ID", "YOD", "DOD", "HOD", "DOB", "Age", "Sex", "Race", "Base_cause", "LineA", "LineB", "LineC", "LineD", "LineII", "Bairro")

#MERGE

homs_01_17 <- do.call(rbind, list(homs_01_10, homs_11_17))
```

Detect and delete duplicates


Then get rid of duplicates
Will compare all fields except potential_ID
homs_01_17

```{r}
df <- homs_01_17@data
df$coords.x1 <- round(homs_01_17@coords[,1], 1)
df$coords.x2 <- round(homs_01_17@coords[,2], 1)

homs_01_17 <- homs_01_17[!duplicated(df),]
rm(df)
```


Creating dates

```{r}
homs_01_17$DOD = as.Date(strptime(homs_01_17$DOD, "%d%m%Y"))
homs_01_17$DOB = as.Date(strptime(homs_01_17$DOB, "%d%m%Y"))
homs_01_17$Age <- interval(homs_01_17$DOB, homs_01_17$DOD) / years(1)
```

Group ages

```{r}
homs_01_17$AgeGroup <- cut(homs_01_17$Age, breaks=c(0, 15, 25, 45, 65, Inf), labels=c("0-14","15-24","25-44","45-64","65+"))
```

##Important Note
There are two shapefiles. One of CTs obtained from IBGE, and the other for Fortaleza at the bairro level obtained from the MHS. The CTs had some mislabelling in terms of the bairros, which could have led to incorrect population denominators for an area. We are using the corrected file.

Now, CTs and bairros are not completely ligned up with each other. So, sometimes a CT in one bairro will jut into another bairro (bairro being defined by the MHS shapefile). However, since we expect to do our analysis at two levels - CT and bairro, it is best to be consistent. So instead of using the MHS shapefile we will use the IBGE shapefile, and create bairros by unions of the nested CTs. But we will use the names used in the Bairro shapefile.

```{r}
shp_CT <- readOGR("Shapefiles/CTs/", "Corrected_CTs")
shp_bairro <- readOGR("Shapefiles/Bairro/", "City_shape_corrected")

#Project CT shapefile into the same projection as point data
def_CRS <- homs_01_17@proj4string
shp_CT <- spTransform(shp_CT, def_CRS)

#Add proper names from bairro shp and get rid of unnecessary columns. Rename to bairro
shp_CT@data <- left_join(shp_CT@data, shp_bairro@data, by=c("CD_GEOCODB"="cod_ibge"))
shp_CT@data <- shp_CT@data[,-14]
colnames(shp_CT@data)[14] <- "Bairro"

shp_CT@data[] <- lapply(shp_CT@data, as.character)

#Dissolving CTs within bairros creates spatialpolygons. TO make spatialpolygonsdataframe, we keep the df from the bairro shapefile and add it afterwards.

df <- shp_bairro@data
rownames(df) <- df$cod_ibge
shp_bairro <- gUnaryUnion(shp_CT, shp_CT@data$CD_GEOCODB)
shp_bairro <- SpatialPolygonsDataFrame(shp_bairro, df)

rm(df)

```

Going to save the new bairro shapefile

```{r}
shp_CT@data[] <- lapply(shp_CT@data, as.character)
writeOGR(shp_bairro, "Shapefiles/Bairro/", "Bairros_from_CTs", driver = "ESRI Shapefile")
```



Merge points and shapes and write files

```{r}
#First merge with CTs. Get rid of unnecessary columns
homs_wth_bairro_CT_df <- cbind(homs_01_17@data[,-15], over(homs_01_17, shp_CT))
homs_wth_bairro_CT_df <- homs_wth_bairro_CT_df[,-16]

#Make this the df for a shapefile
homs_wth_bairro_CT_shp <- homs_01_17
homs_wth_bairro_CT_shp@data <- homs_wth_bairro_CT_df

#Write
write.csv(homs_wth_bairro_CT_df, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/homs_wth_bairro_CT_df.csv")

writeOGR(homs_wth_bairro_CT_shp, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/Shapefiles", "homs_wth_bairro_CT_shp", driver = "ESRI Shapefile")
```

Now I am going to add count information to the CT and bairro shapefiles.

##IMPORTANT NOTE
The total count will still include those with missing age and sex variables.

There are 7 that are not assigned to any bairro.

```{r}
#First for bairros, count using cd_geocodb. Then merge the count data to a df of bairros.

bairros_with_homs <- homs_wth_bairro_CT_df  %>% group_by(CD_GEOCODB) %>% summarise(count = n())

bairros_with_homs <- left_join(shp_bairro@data, bairros_with_homs, by=c("cod_ibge"="CD_GEOCODB"))

colnames(bairros_with_homs) = c("id", "Bairro", "ser", "CD_GEOCODB", "hom_total")

#Next for CTs, count using cd_geocodi. Then merge the count data to a df of CTs
CTs_with_homs <- homs_wth_bairro_CT_df  %>% group_by(CD_GEOCODI) %>% summarise(count = n())

CTs_with_homs <- left_join(shp_CT@data, CTs_with_homs, by="CD_GEOCODI")

#Putting count as zero in the CTs where no homicides were counted.
CTs_with_homs[is.na(CTs_with_homs$count),"count"] <- 0

#Write CSV files
write.csv(bairros_with_homs, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/bairros_with_total_count.csv")
write.csv(CTs_with_homs, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CTs_with_total_count.csv")

```



