---
title: "Spatio-temporal Yearly models"
output: html_notebook
---



```{r}
library(INLA)
library(dplyr)
library(readr)
library(stringi)
library(rgdal)
library(spdep)
```
```{r}
plot_fort_maps2 <- function(data, fill_str, legend_title, plot_title, size=0.1){
  require(ggplot2)
  require(ggsn)
  require(broom)
  ggplot() +                                               # initialize ggplot object
    geom_polygon(                                          # make a polygon
      data = data,                                    # data frame
      aes_string(x = "long", y = "lat", group = "group",                # coordinates, and group them by polygons
                 fill = fill_str),
      size=size, color="black") +                # variable to use for filling
    scale_fill_brewer(name=legend_title, palette = "RdYlBu", direction = -1,
                      drop = FALSE) + # fill with brewer colors   # add title
    theme(line = element_blank(),
          axis.text=element_blank(),      # .. tickmarks..
          axis.title=element_blank(),
          legend.position="none",
          plot.margin = unit(c(0, 0, 0, 0), "cm"),
          #legend.text=element_text(size=15),
          #legend.title=element_text(size=17), # .. axis labels..
          panel.background = element_blank(), plot.title = element_text(size=5)) + ggtitle(plot_title)
}



shape_to_ggplot <- function(shape){
  require(broom)
  gg_data <- tidy(shape)
  data <- slot(shape, "data")
  shape[["polyID"]] <- sapply(slot(shape, "polygons"), function(x) slot(x, "ID"))
  gg_data <- merge(gg_data, shape, by.x="id", by.y="polyID")
  return(gg_data)
}
```


```{r}
SMR_data <- read_csv("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CT_SMR_per_mnth_MISSING_ADJ.csv",
    col_types = cols(CD_GEOCODI = col_character()))
CT_shp <- readOGR("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/Shapefiles/Shapefiles/CTs/", "Corrected_CTs", use_iconv = TRUE, encoding = "latin1")
CT_shp_gg <- shape_to_ggplot(CT_shp)
```

Some CTs have homicides but non pop. We have to revisit these. For now, I am turning Inf IR and SMR to zero. At most there are 2 homicides per year in these CTs. Also NaN are pop zero, hom zero. Also turning these to zero.

```{r}
SMR_data <- SMR_data  %>%
  mutate(Year=cut(Counter, breaks=seq(0,204,12), labels=c(2001:2017))) %>%
  group_by(Year, CD_GEOCODI) %>%
  summarize(y=sum(y), E=sum(E)) %>%
  mutate(SMR=y/E) %>% 
  mutate(SMR=replace(SMR, SMR==Inf | is.nan(SMR), 0)) %>% as.data.frame() %>% rename(obs_count=y, exp_count=E) %>% na.omit()

```

```{r}
CT_shp@data$CD_GEOCODI <- as.character(CT_shp@data$CD_GEOCODI)
CT_shp@data <- arrange(CT_shp@data, CD_GEOCODI)
CT_wm <- poly2nb(CT_shp, snap=0.001)
nb2INLA("CT_graph.adj", CT_wm)
CT_adj <- "CT_graph.adj"
```

Some CTs have homicides but non pop. We have to revisit these. For now, I am turning Inf IR and SMR to zero. At most there are 2 homicides per year in these CTs. Also NaN are pop zero, hom zero. Also turning these to zero. Also turning counts to zero if underlying population is zero

```{r}
yearly_shp_data <- data.frame(CD_GEOCODI=unique(CT_shp@data$CD_GEOCODI)) 

yearly_shp_data <- yearly_shp_data %>% slice(rep(1:n(), each = 17)) %>% arrange(CD_GEOCODI) %>% mutate(bymID=unlist(lapply(1:3044, function(x) rep(x,17))))

yearly_shp_data  <- as.data.frame(yearly_shp_data)

yearly_shp_data$Year <- rep(1:17, 3044)
  
SMR_data$Year <- as.integer(SMR_data$Year)
yearly_shp_data$CD_GECODI <- as.integer(yearly_shp_data$CD_GEOCODI)
  
BYM_data <- SMR_data %>% rename(y=obs_count, E=exp_count) %>% right_join(yearly_shp_data, by=c("CD_GEOCODI", "Year")) %>% mutate(y=replace(y, is.na(y), 0), E=replace(E, is.na(E), 0)) %>% dplyr::select(CD_GEOCODI, y, E, Year) %>% arrange(CD_GEOCODI) %>% mutate(bymID=unlist(lapply(1:3044, function(x) rep(x,17))))

BYM_data$fix_2016 <- ifelse(BYM_data$Year==16,1,0)

BYM_data$year_int <- BYM_data$Year - 1

BYM_data$bymID2 <- BYM_data$bymID
BYM_data$bymID3 <- BYM_data$bymID
```

```{r}
formula_yrlytrnd_2016fix <- y ~ 1 + f(bymID, model="bym",graph=CT_adj, constr=TRUE) + f(bymID2, year_int, model="iid", constr=TRUE) + year_int + f(bymID3, fix_2016, model="iid", constr=TRUE) + fix_2016

model_yrlytrnd_2016fix <- inla(formula_yrlytrnd_2016fix, family="poisson", data=BYM_data, E=E, verbose=TRUE, control.compute=list(dic=TRUE), control.predictor=list(compute=TRUE))

```


```{r}
covariates <- read_csv("Census_data/census_covariates_Fortaleza_01_03_19.csv", 
    col_types = cols(Cod_setor = col_character()))
```

Calculate LII

```{r}

covariates <- covariates %>% mutate(Total_no_HH_in_CT=replace(Total_no_HH_in_CT, is.na(Total_no_HH_in_CT), 0), Total_HH_inc_in_CT=replace(Total_HH_inc_in_CT, is.na(Total_HH_inc_in_CT), 0), Mean_HH_inc=replace(Mean_HH_inc, is.na(Mean_HH_inc), 0))
                                    
                                    
covariates$LII <- NA

for (i in 1:3044){
  CT_Code <- CT_shp@data$CD_GEOCODI[i]
  neighbors_codes <- CT_shp@data$CD_GEOCODI[CT_wm[[i]]]
  total_nb_inc <- sum(covariates[(covariates$Cod_setor %in% neighbors_codes),]$Total_HH_inc_in_CT)
  total_nb_hh <- sum(covariates[(covariates$Cod_setor %in% neighbors_codes),]$Total_no_HH_in_CT)
  Mean_nb_inc <- total_nb_inc / total_nb_hh
  LII <- Mean_nb_inc / covariates[(covariates$Cod_setor)==CT_Code,]$Mean_HH_inc
  covariates[(covariates$Cod_setor==CT_Code),]$LII <- LII
}

covariates$LII[covariates$LII==Inf] <- NA 
covariates$log_LII <- log(covariates$LII)

```

```{r}
BYM_data_covar <- BYM_data %>% left_join(covariates, by=c("CD_GEOCODI"="Cod_setor")) %>% mutate(log_mean_HH_inc=log(Mean_HH_inc)) %>% arrange(CD_GEOCODI) 

BYM_data_covar <- BYM_data_covar %>% mutate(log_mean_HH_inc=replace(log_mean_HH_inc, log_mean_HH_inc==-Inf | is.na(log_mean_HH_inc), 0), log_LII=replace(log_LII, log_LII==-Inf | is.na(log_LII), 0), Perc_Wtr_Spply_Ntwrk=replace(Perc_Wtr_Spply_Ntwrk, is.na(Perc_Wtr_Spply_Ntwrk), 0), Perc_branca=replace(Perc_branca, is.na(Perc_branca), 0), Perc_Garbage_Col_Serv=replace(Perc_Garbage_Col_Serv, is.na(Perc_Garbage_Col_Serv), 0), Perc_PPH_Elec=replace(Perc_PPH_Elec, is.na(Perc_PPH_Elec), 0), Lit_rate=replace(Lit_rate, is.na(Lit_rate), 0), ICE=replace(ICE, is.na(ICE), 0)) 

BYM_data_covar <- BYM_data_covar %>% mutate(ICE_quant=cut(ICE, breaks=quantile(BYM_data_covar$ICE, probs=c(seq(0,1,0.25),1)[1:5]), include.lowest=TRUE))

BYM_data_covar <- BYM_data_covar %>% mutate(log_inc_quant=cut(log_mean_HH_inc, breaks=quantile(BYM_data_covar$log_mean_HH_inc, probs=c(seq(0,1,0.25),1)[1:5]), include.lowest=TRUE))

BYM_data_covar <- BYM_data_covar %>% mutate(log_LII_quant=cut(log_LII, breaks=quantile(BYM_data_covar$log_LII, probs=c(seq(0,1,0.25),1)[1:5]), include.lowest=TRUE))

BYM_data_covar <- BYM_data_covar %>% mutate(log_inc_demean=log_mean_HH_inc-mean(BYM_data_covar$log_mean_HH_inc, na.rm=TRUE))

BYM_data_covar <- BYM_data_covar %>% mutate(Perc_branca_demean=Perc_branca-mean(BYM_data_covar$Perc_branca, na.rm=TRUE))

BYM_data_covar <- BYM_data_covar %>% mutate(Lit_rate_demean=Lit_rate-mean(BYM_data_covar$Lit_rate, na.rm=TRUE))


```

```{r}
formula_yrlytrnd_2016fix_covar <- y ~ 1 + f(bymID, model="bym",graph=CT_adj, constr=TRUE) + f(bymID2, year_int, model="iid", constr=TRUE) + year_int + f(bymID3, fix_2016, model="iid", constr=TRUE) + fix_2016 + log_inc_quant + ICE_quant + log_LII_quant + Lit_rate_demean + Perc_branca_demean

model_yrlytrnd_2016fix_covar <- inla(formula_yrlytrnd_2016fix_covar, family="poisson", data=BYM_data_covar, E=E, verbose=TRUE, control.compute=list(dic=TRUE), control.predictor=list(compute=TRUE))

```

```{r}
formula_only_covar <- y ~ 1 + log_inc_quant + ICE_quant + log_LII_quant + Lit_rate_demean + Perc_branca_demean

model_only_covar<- inla(formula_only_covar, family="poisson", data=BYM_data_covar, E=E, verbose=TRUE, control.compute=list(dic=TRUE), control.predictor=list(compute=TRUE))

```

```{r}
formula_only_covar3 <- y ~ 1 + log_inc_quant + log_LII_quant

model_only_covar3 <- inla(formula_only_covar3, family="poisson", data=BYM_data_covar, E=E, verbose=TRUE, control.compute=list(dic=TRUE), control.predictor=list(compute=TRUE))

formula_only_covar4 <- y ~ 1 + ICE_quant

model_only_covar4 <- inla(formula_only_covar4, family="poisson", data=BYM_data_covar, E=E, verbose=TRUE, control.compute=list(dic=TRUE), control.predictor=list(compute=TRUE))

```

```{r}
formula_only_covar3 <- y ~ 1 + ICE_quant + log_LII_quant + Lit_rate_demean + Perc_branca_demean

model_only_covar3 <- inla(formula_only_covar2, family="poisson", data=BYM_data_covar, E=E, verbose=TRUE, control.compute=list(dic=TRUE), control.predictor=list(compute=TRUE))

```


```{r}
results_null <- data.frame(bymID = BYM_data$bymID, year=BYM_data$Year, smr_fit=model_yrlytrnd_2016fix$summary.fitted.values$mean, global_trnd=rep(model_yrlytrnd_2016fix$summary.fixed$mean[2], 51748), global_2016=rep(model_yrlytrnd_2016fix$summary.fixed$mean[3], 51748))
```

```{r}
results_null$fix_2016 <- ifelse(results_null$year==16,1,0)

results_null$trendfit <- exp((results_null$global_trnd * results_null$year) + (results_null$global_2016 * results_null$fix_2016))

```

```{r}
results_null_plot1 <- dplyr::select(results_null, bymID, year, smr_fit) %>% mutate(type="CT_specific")
results_null_plot2 <- results_null %>% dplyr::select(year, trendfit) %>% rename(smr_fit=trendfit) %>% mutate(type="global") %>% unique() %>% mutate(bymID=3055)
results_null_plot <- rbind(results_null_plot1, results_null_plot2)
```

```{r}
library(ggplot2)
library(plotly)
p <- ggplot() + geom_line(data=results_null_plot, aes(x=year, y=smr_fit, group=bymID, color=type, linetype=type))
ggplotly(p)
```

```{r}
plot_dfs <- lapply(1:17, function(x) filter(select(results_null, year, bymID, smr_fit), year==x))
breaks=c(-1, 0.5, 0.75, 1.0, 1.25, 1.50, 1.75, 2.0, 5.0, Inf)
labels=c("<0.5", "0.5-0.75",
               "0.75-1.0", "1.0-1.25","1.25-1.50","1.5-1.75", "1.75-2.0",
               "2.0-5.0", ">5.0")
plot_dfs <- lapply(plot_dfs, function(x) cbind(x, smr_cut=cut(x$smr_fit, breaks, labels)))
```

```{r}
CT_shp_gg <- left_join(CT_shp_gg, select(BYM_data, bymID, CD_GEOCODI), by="CD_GEOCODI")
plot_dfs <- lapply(plot_dfs, function(x) left_join(CT_shp_gg, x, by="bymID"))
```

```{r}
fit_plots <- vector("list", 17)
for (i in 1:length(plot_dfs)){
  p <- plot_fort_maps2(plot_dfs[[i]], fill_str="smr_cut", legend_title = paste0("Standardized Mortality\nRatio - ", (2000+i)), plot_title = as.character(2000+i))
  print(p)
  fit_plots[[i]] <- p + theme(legend.position = "right", plot.title = element_text(size=22, hjust=0.5))
}
```



```{r}
dir.create("examples")
setwd("examples")

# example 1: simple animated countdown from 10 to "GO!".
png(file="example%02d.png", height=200, width=300, units="mm", res=300)
  for (i in 1:17){
    print(fit_plots[[i]])
  }
dev.off()

# convert the .png files to one .gif file using ImageMagick. 
# The system() function executes the command as if it was done
# in the terminal. the -delay flag sets the time between showing
# the frames, i.e. the speed of the animation.
system("convert -delay 80 *.png example_1.gif")

# to not leave the directory with the single jpeg files
# I remove them.
file.remove(list.files(pattern=".png"))
```

```{r}
mat.marg <- matrix(NA, nrow=51748, ncol=100000)
m <- model_only_covar$marginals.random$ID
for (i in 1:51748){
   u <- m[[51748+i]]
   mat.marg[i,] <- inla.rmarginal(100000, u)
 }
var.u <- apply(mat.marg, 2, var)
We also extract the expected value of the variance for the unstructured component
> var.v <- inla.rmarginal(100000,inla.tmarginal(function(x) 1/x, mod.suicides$marginals.hyper$"Precision for ID (iid component)"))
and  nally we compute the spatial fractional variance fracspatial as
> perc.var.u <- mean(var.u/(var.u+var.v))
> perc.var.u
[1] 0.9418849
```

