---
title: "Counting homicides by age-sex groups and generating SMRs"
output: html_notebook
---

Load libraries

```{r message=FALSE}
library(rgdal)
library(spdep)
library(dplyr)
library(lubridate)
library(maptools)
library(rgeos)
library(readr)
```

Read in required files

```{r}
homs <- read_csv("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/homs_wth_bairro_CT_df.csv",
col_types = cols(CD_GEOCODB = col_character(),
CD_GEOCODD = col_character(), CD_GEOCODI = col_character(),
CD_GEOCODM = col_character(), CD_GEOCODS = col_character(),
DOB = col_date(format = "%Y-%m-%d"),
DOD = col_date(format = "%Y-%m-%d"),
X1 = col_skip()))
```

Count by age, sex and age-sex group

##IMPORTANT NOTE
Will discard those that do not have age or sex information.
There are 159 missing age information due to missing birth date.
There are none missing sex information

For bairros
```{r}
#Keep only relevant columns
bairros_with_homs_by_age_sex <- select(homs, Age, AgeGroup, Sex, CD_GEOCODB)

#Remove those without sex or age info
bairros_with_homs_by_age_sex <- bairros_with_homs_by_age_sex[!is.na(bairros_with_homs_by_age_sex$AgeGroup),]

bairros_with_homs_by_age_sex <- bairros_with_homs_by_age_sex[!is.na(bairros_with_homs_by_age_sex$Sex),]

#Group and count by age sex and bairro code
bairros_with_homs_by_age_sex <- bairros_with_homs_by_age_sex %>% group_by(CD_GEOCODB, AgeGroup, Sex) %>% summarise(hom_count_age_sex = n())

#Now going to create a mock df with all strata. In the previous line strata without observations are missing. After merging, will give these zero count.
df <- as.data.frame(matrix(ncol=0,nrow=1190))
df$Sex <- rep(c(rep("M",5),rep("F",5)),119)
df$AgeGroup <- rep(rep(c("0-14","15-24","25-44","45-64","65+"), 2), 119)
df$CD_GEOCODB <- unlist(lapply(unique(bairros_with_homs_by_age_sex$CD_GEOCODB[!is.na(bairros_with_homs_by_age_sex$CD_GEOCODB)]), function(x) rep(x,10)))

bairros_with_homs_by_age_sex <- left_join(df, bairros_with_homs_by_age_sex, by=c("CD_GEOCODB", "Sex", "AgeGroup"))

bairros_with_homs_by_age_sex$hom_count_age_sex[is.na(bairros_with_homs_by_age_sex$hom_count_age_sex)] <- 0

rm(df)

```

Now for bairros, with age-sex-year strata
```{r}
#Keep only relevant columns
bairros_with_homs_by_age_sex_year <- select(homs, Age, AgeGroup, Sex, CD_GEOCODB, YOD)

#Remove those without sex, age or info
bairros_with_homs_by_age_sex_year <- bairros_with_homs_by_age_sex_year[!is.na(bairros_with_homs_by_age_sex_year$AgeGroup),]

bairros_with_homs_by_age_sex_year <- bairros_with_homs_by_age_sex_year[!is.na(bairros_with_homs_by_age_sex_year$Sex),]

#Group and count by age sex and bairro code
bairros_with_homs_by_age_sex_year <- bairros_with_homs_by_age_sex_year %>% group_by(CD_GEOCODB, AgeGroup, Sex, YOD) %>% summarise(hom_count_age_sex_year = n())

#Now going to create a mock df with all strata. In the previous line strata without observations are missing. After merging, will give these zero count.
df <- as.data.frame(matrix(ncol=0,nrow=20230))
df$Sex <- rep(rep(c(rep("M",5),rep("F",5)), 119), 17)
df$AgeGroup <- rep(rep(rep(c("0-14","15-24","25-44","45-64","65+"), 2), 119), 17)
df$YOD <- rep(unlist(lapply(2001:2017, function(x) rep(x,10))),119)
df$CD_GEOCODB <- unlist(lapply(unique(bairros_with_homs_by_age_sex_year$CD_GEOCODB[!is.na(bairros_with_homs_by_age_sex_year$CD_GEOCODB)]), function(x) rep(x,170)))


bairros_with_homs_by_age_sex_year <- left_join(df, bairros_with_homs_by_age_sex_year, by=c("CD_GEOCODB", "Sex", "AgeGroup", "YOD"))

bairros_with_homs_by_age_sex_year$hom_count_age_sex_year[is.na(bairros_with_homs_by_age_sex_year$hom_count_age_sex_year)] <- 0

rm(df)

```

For CTs

```{r}
#Keep only relevant columns
CTs_with_homs_by_age_sex <- select(homs, Age, AgeGroup, Sex, CD_GEOCODI)

#Remove those without sex or age info
CTs_with_homs_by_age_sex <- CTs_with_homs_by_age_sex[!is.na(CTs_with_homs_by_age_sex$AgeGroup),]

CTs_with_homs_by_age_sex <- CTs_with_homs_by_age_sex[!is.na(CTs_with_homs_by_age_sex$Sex),]

#Group and count by age sex and CTcode
CTs_with_homs_by_age_sex <- CTs_with_homs_by_age_sex %>% group_by(CD_GEOCODI, AgeGroup, Sex) %>% summarise(hom_count_age_sex = n())

#Now going to create a mock df with all strata. In the previous line strata without observations are missing. After merging, will give these zero count.
df <- as.data.frame(matrix(ncol=0,nrow=30440))

#To get unique list of CD_GEOCODI including those without homicides we need to read in a different file
CTs_with_total_count <- read_csv("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CTs_with_total_count.csv", col_types = cols(CD_GEOCODI = col_character(), ID = col_skip(), X1 = col_skip()))

df$CD_GEOCODI <- unlist(lapply(unique(CTs_with_total_count$CD_GEOCODI), function(x) rep(x,10)))
rm(CTs_with_total_count)

df$Sex <- rep(c(rep("M",5),rep("F",5)),3044)
df$AgeGroup <- rep(rep(c("0-14","15-24","25-44","45-64","65+"), 2), 3044)

CTs_with_homs_by_age_sex <- left_join(df, CTs_with_homs_by_age_sex, by=c("CD_GEOCODI", "Sex", "AgeGroup"))

CTs_with_homs_by_age_sex$hom_count_age_sex[is.na(CTs_with_homs_by_age_sex$hom_count_age_sex)] <- 0

rm(df)

```

For CTs with Years

```{r}
#Keep only relevant columns
CTs_with_homs_by_age_sex_year <- select(homs, Age, AgeGroup, Sex, CD_GEOCODI, YOD)

#Remove those without sex or age info
CTs_with_homs_by_age_sex_year <- CTs_with_homs_by_age_sex_year[!is.na(CTs_with_homs_by_age_sex_year$AgeGroup),]

CTs_with_homs_by_age_sex_year <- CTs_with_homs_by_age_sex_year[!is.na(CTs_with_homs_by_age_sex_year$Sex),]

#Group and count by age sex and CTcode
CTs_with_homs_by_age_sex_year <- CTs_with_homs_by_age_sex_year %>% group_by(CD_GEOCODI, AgeGroup, Sex, YOD) %>% summarise(hom_count_age_sex_year = n())

#Now going to create a mock df with all strata. In the previous line strata without observations are missing. After merging, will give these zero count.
df <- as.data.frame(matrix(ncol=0,nrow=517480))

#To get unique list of CD_GEOCODI including those without homicides we need to read in a different file
CTs_with_total_count <- read_csv("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CTs_with_total_count.csv", col_types = cols(CD_GEOCODI = col_character(), ID = col_skip(), X1 = col_skip()))


df$CD_GEOCODI <- unlist(lapply(unique(CTs_with_total_count$CD_GEOCODI), function(x) rep(x,170)))

rm(CTs_with_total_count)

df$Sex <- rep(rep(c(rep("M",5),rep("F",5)),3044), 17)
df$AgeGroup <- rep(rep(rep(c("0-14","15-24","25-44","45-64","65+"), 2), 3044), 17)
df$YOD <- rep(unlist(lapply(2001:2017, function(x) rep(x,10))),3044)

CTs_with_homs_by_age_sex_year <- left_join(df, CTs_with_homs_by_age_sex_year, by=c("CD_GEOCODI", "Sex", "AgeGroup", "YOD"))

CTs_with_homs_by_age_sex_year$hom_count_age_sex_year[is.na(CTs_with_homs_by_age_sex_year$hom_count_age_sex_year)] <- 0

rm(df)

```

Now I have to get expected counts. For this I have to get city-wide rates by age-sex-year strata

Reading in stratified population data for whole city
```{r}
pop_city <- read_csv("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit/Census_data/Fortaleza_Age_Sex_Stratified_pop.csv")[,-1]
pop_city$Sex <- c(rep("M",5),rep("F",5))
pop_city$AgeGroup <- rep(c("0-14","15-24","25-44","45-64","65+"), 2)

#We are going to collapse all data from the bairro file to create homicide counts for each age group and sex
city_homs <- bairros_with_homs_by_age_sex %>% group_by(AgeGroup, Sex) %>% summarise(hom_count = sum(hom_count_age_sex))

city_rates <- left_join(city_homs, pop_city, by=c("Sex", "AgeGroup"))
city_rates$rate17yr <- city_rates$hom_count / city_rates$Pop
city_rates$rate_per_yr <- city_rates$rate17yr/17
```

Now by applying these rates to census tract age and sex groups we can find expected count of homicides
```{r}
pop_ct <- read_csv("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit/Census_data/Fortaleza_by_CT_Age_Sex_Stratified_pop.csv", 
    col_types = cols(Cod_setor = col_character()))[,-1]

pop_ct$Sex <- rep(c(rep("M",5),rep("F",5)),3021)
pop_ct$AgeGroup <- rep(rep(c("0-14","15-24","25-44","45-64","65+"), 2), 3021)

ct_expected_counts <- left_join(pop_ct, select(city_rates, Sex, AgeGroup, rate17yr, rate_per_yr), by=c("Sex", "AgeGroup"))

```

```{r}
ct_expected_counts$exp_count_17yr <- ct_expected_counts$population * ct_expected_counts$rate17yr
ct_expected_counts$exp_count_per_yr <- ct_expected_counts$population * ct_expected_counts$rate_per_yr
colnames(ct_expected_counts)[1] <- "CD_GEOCODI"
```

Now to get actual number of homicides per census tract and create SMR in aggregate for all 17 years

```{r}
ct_SMR_17yr <- left_join(ct_expected_counts, CTs_with_homs_by_age_sex, by=c("Sex", "AgeGroup", "CD_GEOCODI"))
colnames(ct_SMR_17yr)[10] <- "obs_count_age_sex"
ct_SMR_17yr$obs_count_age_sex[is.na(ct_SMR_17yr$obs_count_age_sex)] <- 0
ct_SMR_17yr$exp_count_17yr[is.na(ct_SMR_17yr$exp_count_17yr)] <- 0
ct_SMR_17yr$population[is.na(ct_SMR_17yr$population)] <- 0

ct_SMR_17yr <- ct_SMR_17yr %>% group_by(CD_GEOCODI) %>% summarise(obs_count = sum(obs_count_age_sex), exp_count = sum(exp_count_17yr), pop=sum(population))

ct_SMR_17yr$SMR_17yr <- ct_SMR_17yr$obs_count / ct_SMR_17yr$exp_count
ct_SMR_17yr$IR_17yr <- (ct_SMR_17yr$obs_count / (ct_SMR_17yr$pop*17)) * 100000
```

Now to get SMR for each year. Note that we do not create expected rates for each year specifically. We use mean from the entire period.

```{r}
ct_SMR_per_yr <- left_join(ct_expected_counts, CTs_with_homs_by_age_sex_year, by=c("Sex", "AgeGroup", "CD_GEOCODI"))

colnames(ct_SMR_per_yr)[11] <- "obs_count_age_sex"

ct_SMR_per_yr$obs_count_age_sex[is.na(ct_SMR_per_yr$obs_count_age_sex)] <- 0
ct_SMR_per_yr$exp_count_per_yr[is.na(ct_SMR_per_yr$exp_count_per_yr)] <- 0
ct_SMR_per_yr$population[is.na(ct_SMR_per_yr$population)] <- 0

ct_SMR_per_yr <- ct_SMR_per_yr %>% group_by(CD_GEOCODI, YOD) %>% summarise(obs_count = sum(obs_count_age_sex), exp_count = sum(exp_count_per_yr), pop=sum(population))

ct_SMR_per_yr$SMR_per_yr <- ct_SMR_per_yr$obs_count / ct_SMR_per_yr$exp_count
ct_SMR_per_yr$IR_per_yr <- (ct_SMR_per_yr$obs_count / ct_SMR_per_yr$pop)*100000

```

Now I am going to do the same thing (generate aggregate and yearly SMR and IR) at the bairro level as well.

##NOTHING UNDER THIS LINE HAS BEEN CHANGED FROM CT TO BAIRRO YET.

```{r}
#First going convert CT population to bairro population
CTs_to_bairro <- distinct(select(homs, CD_GEOCODI, CD_GEOCODB))
CTs_to_bairro <- CTs_to_bairro[complete.cases(CTs_to_bairro),]

pop_bairro <- left_join(pop_ct, CTs_to_bairro, by=c("Cod_setor"="CD_GEOCODI")) %>% group_by(CD_GEOCODB, AgeGroup, Sex) %>% summarize(population=sum(population, na.rm=TRUE))
#pop_bairro[is.na(pop_bairro$population),]$population <- 0
pop_bairro <- pop_bairro[complete.cases(pop_bairro),]

bairro_expected_counts <- left_join(pop_bairro, select(city_rates, Sex, AgeGroup, rate17yr, rate_per_yr), by=c("Sex", "AgeGroup"))

```

```{r}
bairro_expected_counts$exp_count_17yr <- bairro_expected_counts$population * bairro_expected_counts$rate17yr
bairro_expected_counts$exp_count_per_yr <- bairro_expected_counts$population * bairro_expected_counts$rate_per_yr
```

Now to get actual number of homicides per bairro and create SMR in aggregate for all 17 years

```{r}
bairro_SMR_17yr <- left_join(bairro_expected_counts, bairros_with_homs_by_age_sex, by=c("Sex", "AgeGroup", "CD_GEOCODB"))
colnames(bairro_SMR_17yr)[9] <- "obs_count_age_sex"
bairro_SMR_17yr$obs_count_age_sex[is.na(bairro_SMR_17yr$obs_count_age_sex)] <- 0
bairro_SMR_17yr$exp_count_17yr[is.na(bairro_SMR_17yr$exp_count_17yr)] <- 0
bairro_SMR_17yr$population[is.na(bairro_SMR_17yr$population)] <- 0

bairro_SMR_17yr <- bairro_SMR_17yr %>% group_by(CD_GEOCODB) %>% summarise(obs_count = sum(obs_count_age_sex), exp_count = sum(exp_count_17yr), pop=sum(population))

bairro_SMR_17yr$SMR_17yr <- bairro_SMR_17yr$obs_count / bairro_SMR_17yr$exp_count
bairro_SMR_17yr$IR_17yr <- (bairro_SMR_17yr$obs_count / (bairro_SMR_17yr$pop*17)) * 100000
```

Now to get SMR for each year. Note that we do not create expected rates for each year specifically. We use mean from the entire period.

```{r}
bairro_SMR_per_yr <- left_join(bairro_expected_counts, bairros_with_homs_by_age_sex_year, by=c("Sex", "AgeGroup", "CD_GEOCODB"))

colnames(bairro_SMR_per_yr)[10] <- "obs_count_age_sex"

bairro_SMR_per_yr$obs_count_age_sex[is.na(bairro_SMR_per_yr$obs_count_age_sex)] <- 0
bairro_SMR_per_yr$exp_count_per_yr[is.na(bairro_SMR_per_yr$exp_count_per_yr)] <- 0
bairro_SMR_per_yr$population[is.na(bairro_SMR_per_yr$population)] <- 0

bairro_SMR_per_yr <- bairro_SMR_per_yr %>% group_by(CD_GEOCODB, YOD) %>% summarise(obs_count = sum(obs_count_age_sex), exp_count = sum(exp_count_per_yr), pop=sum(population))

bairro_SMR_per_yr$SMR_per_yr <- bairro_SMR_per_yr$obs_count / bairro_SMR_per_yr$exp_count
bairro_SMR_per_yr$IR_per_yr <- (bairro_SMR_per_yr$obs_count / bairro_SMR_per_yr$pop)*100000

```


For Bairros by Months

```{r}
homs$MOD <- substr(homs$DOD, 6,7)
homs$Counter <- (as.integer(homs$YOD)-2001)*12 + as.integer(homs$MOD)



#Keep only relevant columns
bairros_with_homs_by_age_sex_mnth <- select(homs, Age, AgeGroup, Sex, CD_GEOCODB, Counter, YOD, MOD)

#Remove those without sex, age or info
bairros_with_homs_by_age_sex_mnth <- bairros_with_homs_by_age_sex_mnth[!is.na(bairros_with_homs_by_age_sex_mnth$AgeGroup),]

bairros_with_homs_by_age_sex_mnth <- bairros_with_homs_by_age_sex_mnth[!is.na(bairros_with_homs_by_age_sex_mnth$Sex),]

#Group and count by age sex and bairro code
bairros_with_homs_by_age_sex_mnth <- bairros_with_homs_by_age_sex_mnth %>% group_by(CD_GEOCODB, AgeGroup, Sex, Counter) %>% summarise(hom_count_age_sex_mnth = n(), YOD = mean(as.integer(YOD)), MOD=mean(as.integer(MOD)))

#Now going to create a mock df with all strata. In the previous line strata without observations are missing. After merging, will give these zero count.
df <- as.data.frame(matrix(ncol=0,nrow=242760))
df$Sex <- rep(rep(rep(c(rep("M",5),rep("F",5)), 119), 12), 17)
df$AgeGroup <- rep(rep(rep(rep(c("0-14","15-24","25-44","45-64","65+"), 2), 119), 12), 17)
df$Counter <- rep(unlist(lapply(1:204, function(x) rep(x,10))),119)
df$CD_GEOCODB <- unlist(lapply(unique(bairros_with_homs_by_age_sex_mnth$CD_GEOCODB[!is.na(bairros_with_homs_by_age_sex_mnth$CD_GEOCODB)]), function(x) rep(x,2040)))


bairros_with_homs_by_age_sex_mnth <- left_join(df, bairros_with_homs_by_age_sex_mnth, by=c("CD_GEOCODB", "Sex", "AgeGroup", "Counter"))

bairros_with_homs_by_age_sex_mnth$hom_count_age_sex_mnth[is.na(bairros_with_homs_by_age_sex_mnth$hom_count_age_sex_mnth)] <- 0

rm(df)

```

```{r}

bairro_SMR_per_mnth <- left_join(bairro_expected_counts, bairros_with_homs_by_age_sex_mnth, by=c("Sex", "AgeGroup", "CD_GEOCODB"))
colnames(bairro_SMR_per_mnth)[10] <- "obs_count_age_sex"

bairro_SMR_per_mnth$exp_count_per_mnth <- bairro_SMR_per_mnth$exp_count_per_yr / 12

bairro_SMR_per_mnth$obs_count_age_sex[is.na(bairro_SMR_per_mnth$obs_count_age_sex)] <- 0
bairro_SMR_per_mnth$exp_count_per_mnth[is.na(bairro_SMR_per_mnth$exp_count_per_mnth)] <- 0
bairro_SMR_per_mnth$population[is.na(bairro_SMR_per_mnth$population)] <- 0

bairro_SMR_per_mnth <- bairro_SMR_per_mnth %>% group_by(CD_GEOCODB, Counter) %>% summarise(obs_count = sum(obs_count_age_sex), exp_count = sum(exp_count_per_mnth), pop=sum(population))

bairro_SMR_per_mnth$SMR_per_mnth <- bairro_SMR_per_mnth$obs_count / bairro_SMR_per_mnth$exp_count
bairro_SMR_per_mnth$IR_per_mnth <- (bairro_SMR_per_mnth$obs_count / bairro_SMR_per_mnth$pop)*100000

rm(df)

```


```{r}
homs$MOD <- substr(homs$DOD, 6,7)
homs$Counter <- (as.integer(homs$YOD)-2001)*12 + as.integer(homs$MOD)



#Keep only relevant columns
bairros_with_homs_by_age_sex_mnth <- select(homs, Age, AgeGroup, Sex, CD_GEOCODB, Counter, YOD, MOD)

#Remove those without sex, age or info
bairros_with_homs_by_age_sex_mnth <- bairros_with_homs_by_age_sex_mnth[!is.na(bairros_with_homs_by_age_sex_mnth$AgeGroup),]

bairros_with_homs_by_age_sex_mnth <- bairros_with_homs_by_age_sex_mnth[!is.na(bairros_with_homs_by_age_sex_mnth$Sex),]

#Group and count by age sex and bairro code
bairros_with_homs_by_age_sex_mnth <- bairros_with_homs_by_age_sex_mnth %>% group_by(CD_GEOCODB, AgeGroup, Sex, Counter) %>% summarise(hom_count_age_sex_mnth = n(), YOD = mean(as.integer(YOD)), MOD=mean(as.integer(MOD)))

#Now going to create a mock df with all strata. In the previous line strata without observations are missing. After merging, will give these zero count.
df <- as.data.frame(matrix(ncol=0,nrow=242760))
df$Sex <- rep(rep(rep(c(rep("M",5),rep("F",5)), 119), 12), 17)
df$AgeGroup <- rep(rep(rep(rep(c("0-14","15-24","25-44","45-64","65+"), 2), 119), 12), 17)
df$Counter <- rep(unlist(lapply(1:204, function(x) rep(x,10))),119)
df$CD_GEOCODB <- unlist(lapply(unique(bairros_with_homs_by_age_sex_mnth$CD_GEOCODB[!is.na(bairros_with_homs_by_age_sex_mnth$CD_GEOCODB)]), function(x) rep(x,2040)))


bairros_with_homs_by_age_sex_mnth <- left_join(df, bairros_with_homs_by_age_sex_mnth, by=c("CD_GEOCODB", "Sex", "AgeGroup", "Counter"))

bairros_with_homs_by_age_sex_mnth$hom_count_age_sex_mnth[is.na(bairros_with_homs_by_age_sex_mnth$hom_count_age_sex_mnth)] <- 0

rm(df)

```

For CTs per month

```{r}

#Keep only relevant columns
CTs_with_homs_by_age_sex_mnth <- select(homs, Age, AgeGroup, Sex, CD_GEOCODI, Counter, YOD, MOD)

#Remove those without sex, age or info
CTs_with_homs_by_age_sex_mnth <- CTs_with_homs_by_age_sex_mnth[!is.na(CTs_with_homs_by_age_sex_mnth$AgeGroup),]

CTs_with_homs_by_age_sex_mnth <- CTs_with_homs_by_age_sex_mnth[!is.na(CTs_with_homs_by_age_sex_mnth$Sex),]

#Group and count by age sex and bairro code
CTs_with_homs_by_age_sex_mnth <- CTs_with_homs_by_age_sex_mnth %>% group_by(CD_GEOCODI, AgeGroup, Sex, Counter) %>% summarise(hom_count_age_sex_mnth = n(), YOD = mean(as.integer(YOD)), MOD=mean(as.integer(MOD)))

#Now going to create a mock df with all strata. In the previous line strata without observations are missing. After merging, will give these zero count.
df <- as.data.frame(matrix(ncol=0,nrow=6209760))
df$Sex <- rep(rep(rep(c(rep("M",5),rep("F",5)), 3044), 12), 17)
df$AgeGroup <- rep(rep(rep(rep(c("0-14","15-24","25-44","45-64","65+"), 2), 3044), 12), 17)
df$Counter <- rep(unlist(lapply(1:204, function(x) rep(x,10))),3044)

CTs_with_total_count <- read_csv("~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CTs_with_total_count.csv", col_types = cols(CD_GEOCODI = col_character(), ID = col_skip(), X1 = col_skip()))


df$CD_GEOCODI <- unlist(lapply(unique(CTs_with_total_count$CD_GEOCODI), function(x) rep(x,2040)))

rm(CTs_with_total_count)

CTs_with_homs_by_age_sex_mnth <- left_join(df, CTs_with_homs_by_age_sex_mnth, by=c("CD_GEOCODI", "Sex", "AgeGroup", "Counter"))

CTs_with_homs_by_age_sex_mnth$hom_count_age_sex_mnth[is.na(CTs_with_homs_by_age_sex_mnth$hom_count_age_sex_mnth)] <- 0

rm(df)

```

```{r}

ct_SMR_per_mnth <- left_join(ct_expected_counts, CTs_with_homs_by_age_sex_mnth, by=c("Sex", "AgeGroup", "CD_GEOCODI"))
colnames(ct_SMR_per_mnth)[11] <- "obs_count_age_sex"

ct_SMR_per_mnth$exp_count_per_mnth <- ct_SMR_per_mnth$exp_count_per_yr / 12

ct_SMR_per_mnth$obs_count_age_sex[is.na(ct_SMR_per_mnth$obs_count_age_sex)] <- 0
ct_SMR_per_mnth$exp_count_per_mnth[is.na(ct_SMR_per_mnth$exp_count_per_mnth)] <- 0
ct_SMR_per_mnth$population[is.na(ct_SMR_per_mnth$population)] <- 0

ct_SMR_per_mnth <- ct_SMR_per_mnth %>% group_by(CD_GEOCODI, Counter) %>% summarise(obs_count = sum(obs_count_age_sex), exp_count = sum(exp_count_per_mnth), pop=sum(population))

ct_SMR_per_mnth$SMR_per_mnth <- ct_SMR_per_mnth$obs_count / ct_SMR_per_mnth$exp_count
ct_SMR_per_mnth$IR_per_mnth <- (ct_SMR_per_mnth$obs_count / ct_SMR_per_mnth$pop)*100000

rm(df)

```

```{r}
ct_SMR_per_period <- ct_SMR_per_mnth %>% mutate(period=ifelse(Counter<=60, 1, ifelse(Counter>60 & Counter<=120, 2, ifelse(Counter>120 & Counter <=156, 3, ifelse(Counter>156 & Counter <=190, 4, ifelse(Counter>190 & Counter<=204, 5, NA)))))) %>% group_by(CD_GEOCODI, period) %>% summarize(obs_count=sum(obs_count), exp_count=sum(exp_count), pop=mean(pop)) %>% mutate(SMR_per_period=obs_count/exp_count)

```


Now I am going to write all the files.

```{r}
write_csv(bairro_SMR_17yr, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/Bairro_SMR_IR_17yr-agg.csv")
write_csv(bairro_SMR_per_yr, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/Bairro_SMR_IR_per_yr.csv")
write_csv(ct_SMR_17yr, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CT_SMR_IR_17yr-agg.csv")
write_csv(ct_SMR_per_yr, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CT_SMR_IR_per_yr-agg.csv")
write_csv(bairro_SMR_per_mnth, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/Bairro_SMR_IR_per_mnth.csv")
write_csv(ct_SMR_per_mnth, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CT_SMR_IR_per_mnth.csv")
write_csv(ct_SMR_per_period, "~/Documents/Harvard - SM80/Thesis/Fortaleza_Hom_RGit_PRIVATE_Files/CT_SMR_per_period.csv")

```




